---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)

cleaned_data <- read_csv(here::here("outputs/Data/analysis_data.csv"))

```


# Introduction

The introduction is self-contained and tells a reader everything they need to know including: 1) broader context to motivate; 2) some detail about what the paper is about; 3) a clear gap that needs to be filled; 4) what was done; 5) what was found; 6) why it is important; 7) the structure of the paper. A reader should be able to read only the introduction and know what was done, why, and what was found. Likely 3 or 4 paragraphs, or 10 per cent of total.

You can and should cross-reference sections and sub-sections. 

The remainder of this paper is structured as follows. @sec-data....



# Data {#sec-data}

The data used in this paper is from (cite og paper). The data used in the paper is extremely compplicated as it combines numerous data sets to complete their analysis. In this paper, we wanted to simplify the data 
```{r}

ggplot(cleaned_data, aes(x = reorder(hhld_inc, -table(hhld_inc)[hhld_inc]), fill = hhld_inc)) +
  geom_bar() +
  labs(title = "Distribution of Household Income",
       x = "Household Income",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
ggplot(cleaned_data, aes(x = hhld_inc, fill = hhld_inc)) +
  geom_bar() +
  labs(title = "Distribution of Household Income",
       x = "Household Income",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
ggplot(cleaned_data, aes(x = factor(hhld_inc, levels = unique(hhld_inc)), fill = hhld_inc)) +
  geom_bar() +
  labs(title = "Distribution of Household Income",
       x = "Household Income",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```
```{r}
ggplot(cleaned_data, aes(x = factor(hhld_inc, levels = unique(cleaned_data$hhld_inc)), fill = hhld_inc)) +
  geom_bar() +
  labs(title = "Distribution of Household Income",
       x = "Household Income",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```



\newpage


# References


