---
title: ""
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(knitr)
cleaned_data <- read_csv(here::here("outputs/Data/analysis_data.csv"))

```


# Introduction

The introduction is self-contained and tells a reader everything they need to know including: 1) broader context to motivate; 2) some detail about what the paper is about; 3) a clear gap that needs to be filled; 4) what was done; 5) what was found; 6) why it is important; 7) the structure of the paper. A reader should be able to read only the introduction and know what was done, why, and what was found. Likely 3 or 4 paragraphs, or 10 per cent of total.


# Data {#sec-data}
## Data Introduction {#sec-dataintro}
The data used in this paper is from (cite og paper). The data used in the paper is extremely complicated as it combines numerous data sets to complete their analysis. In this paper, we wanted to simplify the data. To simply the original papers data we focused on one of their eleven data sets. The most comprehensive data set was their baseline data. This data was collected through a facebook ad. The respondents answered a number of questions including questions about income, ethnicity, family, political beliefs and political following. The data was cleaned to focus on respondents income, race and how closely they follow politics. This provided a data set that could answer a number of questions. To first understand the possible research questions the data it must to be visualized as it has nearly 6000 entries. 

## Income Data {#sec-income_data}
The first visualization of the data is in @figure1. This table illustrates that the least number of households make greater than 100,000 USD, below 20,000USD or preferred not to answer. This closely resembles the expected distribution of USA household income. As expected in any income distribution the majority of responses fall within the average income ranges of the USA, between 20,000USD and 100,000USD
```{r}
#| label: fig-figure1
#| fig-cap: "Distribution of Income in Responses"
#| warning: false
#| message: false
#| echo: false
ggplot(cleaned_data, aes(x = reorder(hhld_inc, -table(hhld_inc)[hhld_inc]), fill = hhld_inc)) +
  geom_bar() +
  labs(x = "Household Income",
       y = "Number of Respondents") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```
## Ethnicity Data {#sec-race_data}
describe table and its implications and relevence
```{r}
#| label: tbl-race_dist
#| tbl-cap: "Percentage of each Ethnicity Within the Data"
#| warning: false
#| message: false
#| echo: false
proportion_table <- cleaned_data %>%
  group_by(race) %>%
  summarise((proportion = n() / nrow(cleaned_data))*100) 
kable(proportion_table)

```

## Politics Data {#sec-pol_data}


# Model

simple regression of income and politics data controlling for race 

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}



\newpage


# References


